---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# comat

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/Nowosad/comat.svg?branch=master)](https://travis-ci.org/Nowosad/comat)
[![Codecov test coverage](https://codecov.io/gh/Nowosad/comat/branch/master/graph/badge.svg)](https://codecov.io/gh/Nowosad/comat?branch=master)
[![CRAN status](https://www.r-pkg.org/badges/version/comat)](https://CRAN.R-project.org/package=comat)
[![CRAN RStudio mirror downloads](http://cranlogs.r-pkg.org/badges/comat)](https://cran.r-project.org/package=comat)
<!-- badges: end -->

The goal of **comat** is to create co-occurrence matrices based on spatial data, including a weighted co-occurrence matrix (*wecoma*) and an integrated co-occurrence matrix (*incoma*).

## Installation

You can install the released version of **comat** from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("comat")
```

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("Nowosad/comat")
```
## Example

This is a basic example which shows you how to create a weighted co-occurrence matrix (*wecoma*) based on two simple rasters (for simplicity presented as matrices). 
The first one `raster_x` represents some categories, and the second one `raster_w` represents weights.

```{r}
library(comat)
library(raster)
data(raster_x, package = "comat")
data(raster_w, package = "comat")
raster_x
raster_w
```

The `get_wecoma()` function can be next used to create a weighted co-occurrence matrix.

```{r}
get_wecoma(
  raster_x,
  raster_w,
  neighbourhood = 4
)
```

This function allows for some parametrization using additional arguments, e.g.:

```{r}
get_wecoma(
  raster_x,
  raster_w,
  neighbourhood = 4,
  fun = "focal",
  na_action = "keep"
)
```

## Documentation

For more examples see the package's vignettes:

1. [A co-occurrence matrix (coma) representation](https://nowosad.github.io/comat/articles/coma.html)
2. [A weighted co-occurrence matrix (wecoma) representation](https://nowosad.github.io/comat/articles/wecoma.html)
3. [An integrated co-occurrence matrix (incoma) representation](https://nowosad.github.io/comat/articles/incoma.html)

## Contribution

Contributions to this package are welcome.
The preferred method of contribution is through a GitHub pull request. 
Feel free to contact me by creating [an issue](https://github.com/Nowosad/comat/issues). 

## Citation

To cite the `comat` package in publications, please use [this paper](https://doi.org/10.1016/j.apgeog.2020.102239):

Nowosad J, Stepinski TF (2021) Pattern-based identification and mapping of landscape
  types using multi-thematic data, International Journal of Geographical Information
  Science, DOI: 10.1080/13658816.2021.1893324 

LaTeX/BibTeX version can be obtained with:

```
library(comat)
citation("comat")
```
