---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# comat

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/Nowosad/wecoma.svg?branch=master)](https://travis-ci.org/Nowosad/wecoma)
<!-- badges: end -->

The goal of **comat** is to create cooccurrence matrices based on spatial data, including a weighted coocurrence matrix (*wecoma*).

## Installation

<!-- You can install the released version of comat from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ``` r -->
<!-- install.packages("comat") -->
<!-- ``` -->

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("Nowosad/wecoma")
```
## Example

This is a basic example which shows you how to create a weighted cooccurrence matrix based on two simple rasters. 
The first one `x` represents some categories, and the second one `w` represents weights.

```{r}
library(comat)
library(raster)
data(x, package = "comat")
data(w, package = "comat")
par(mfcol = c(1, 2))
plot(x, main = "Categories")
plot(w, main = "Weights")
```

The `get_wecoma` function can be next used to create a weighted cooccurrence matrix.

```{r}
get_wecoma(x, w)
```

This function allows for some parametrization using additional arguments, e.g. :

```{r}
get_wecoma(x, w, fun = "focal", na_action = "keep")
```

